Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression\thanksref{T1}

### 2018-11-03 revision_1/paper.tex
fixes to simulation post-processing code and corresponding edits to the simulation section of paper

To complete the specification of model \eqref{fullsim2},  $a_{s}$ and $b_{s}$ are fixed to a variety of values representing different levels of prior knowledge. For each we set $b_{s} = 4a_{s}c$ resulting in a prior mean for each $\sigma^{2}_{i}$ of $\frac{4ca_{s}}{a_{s}-1}, \ a_{s} >1$. The precision is $\frac{(a_{s} -1)^{2}(a_{s}-2)}{(4ca_{s})^{2}}$; meaning the larger $a_{s}$, the more informative the prior. With $c = 1$ the shrinkage (for large $a_{s}$) is to the true value of $\sigma^{2} = 4$. We consider $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$. %Values of $c \neq 1$ will result in shrinkage towards the wrong value of $\sigma^{2}$.

@@

To assess predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
\begin{equation}
\label{kl}
KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
\end{equation}
where $M$ indexes the fitting model, $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$; the normal density function with mean $\theta_{i}$ and variance $\sigma^{2}$, evaluated at $\tilde y$. For the Bayesian models ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are done separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models, where smaller means correspond to better fits. Sampling variation is summarized with the standard error between the $K=30$ replicates in the simulation:  $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.

Figure \ref{kl_sim} displays $\overline{KL}^{(M)}_{{\cdot}{\cdot}}$  with error-bars plus/minus one $SE(\overline{KL}^{(M)}_{{\cdot} k})$ for each $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$. %The panels are for $c = 0.5, 1, 2$ with $a_{s} = 1.25,  5, 10$ on the abscissa. 
The values of $a_{s}$ and $c$, do not effect the classical robust linear models. The average KL  for the normal theory model ranges between $0.31$ and $0.37$ which is much worse than the robust methods and hence these are left out of the figure. Overall, the results are strikingly in favor of the restricted likelihood methods for the range of hyper-parameter values studied. Undoubtably, the most precise and accurate prior studied is  $c = 1$ and $a_{s} = 10$  and this results in the lowest (and best) average KL for both the Tukey and Huber restricted likelihood versions (shown in the middle panel). $c = 0.5$ performs the worst, but still better than the classical fits. There is evidence that performance starts to degrade in this case when $a_{s} = 10$ as reflected in a larger average KL. Here, the prior mean and precision are $2.22$ and $1.62$ and we suspect this is starting to put too much mass on $\sigma^{2}_{i}$ values much smaller than $\sigma^{2}=4$. For $c = 2$, the performance still improves from $a_{s} = 5$ to $a_{s} = 10$. Here the mean is $8.89$ and precision is only $0.1$; apparently not yet large enough to degrade the performance due to an incorrect mean. Lastly, Tukey's statistic performs marginally better than Huber's. This is likely due to the fact that Tukey's estimator trims extreme outliers completely in the estimation procedure \citep{huber2009}.

@@

%% DELETE?
%It is also interesting to consider the effects of factors $n_{i}$, $p_{i}$, and $m_{i}$. For a given factor and simulation, the $KL^{(M)}_{ik}$  are averaged by factor level. For the Bayesian models, the averages are also taken over the different values of $a_{s}$ and $c$. Figure \ref{kl_mnp} displays these averages for $m,n,$ and $p$ with error bars plus/minus one standard error. The restricted likelihood versions consistently perform better than their classical counterparts. Intuitively, as the amount of contamination ($p$) increases performance degrades as it becomes more difficult to identify the good data. Likewise, as $n$ increases, the performance for the Bayesian methods become closer to that of the their classical counterparts reflecting the diminishing effect of the prior. However, the decrease of KL-diveregence with $m$ and increase with $n$ is somewhat surprising. To investigate, Figure \ref{boxTheta} and \ref{boxSigma} display boxplots of $(\theta_{i} - \hat\theta_{i})$ and $\hat\sigma_{i}$ for each simulated data set. The $\hat\theta_{i}$'s have no systematic bias however the $\hat\sigma_{i}$'s do; consistently biased upward from the value of $\sigma = 2$ for the good portion of data. As $n$ increases, the bias remains the same and the variation in the estimates gets smaller. Thus, the estimates are getting more certain about an incorrect value; explaining the degradation of the KL-divergence. As $m$ increases, the variance in the estimates gets larger and there is a marginal increase in the bias. Thus, for $m=9$ there is more certainty about an incorrect value and  the increased uncertainty at $m=25$ helps to improve the KL-divergence. %Finally, as $p$ grows the bias of the estimates gets significantly larger; resulting in poorer estimation and poorer KL-divergences.
% 
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{kl_sim2_mnp.png}
%\caption{Average KL - divergence plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right)}
%\label{kl_mnp}
%\end{figure}
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{rlm_mnp_theta.png}
%\caption{Boxplots $(\theta_{i} - \hat\theta_{i})$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right) where $\hat\theta_{i}$ are the classical robust estimators (Huber's and Tukey's).}
%\label{boxTheta}
%\end{figure}
%\begin{figure}[t]
%\centering
%\includegraphics[width = 4in]{rlm_mnp_sigma.png}
%\caption{Boxplots of the classical robust estimators (Huber's and Tukey's) for $\sigma_{i}$ across all simulations separated by the values for $m$ (left), $n$ (middle), $p$ (right). The horizontal line at $\sigma = 2$ highlights the true standard deviation of the `good' data.}
%\label{boxSigma}
%\end{figure}
%% --------
{EDIT THIS} This simulation shows the potential of the restricted likelihood while highlighting some interesting observations. Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, the standard choice of $95\%$ efficiency at the normal was used.  Under the data generating model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made in both the classical and Bayesian settings. The Bayesian setting allows for the incorporation of informative prior information which, as shown in this example, can dramatically improve prediction. The results are sensitive to the prior, but here we have observed good relative improvement over the classical counterparts for a range of prior choices.

@@

It is also interesting to examine the results within each state. Figure \ref{fig:hierTLMstate} summarizes ${TLM}_b(A)_{j}$ for each state where the points and errorbars are the averages and plus/minus one standard deviation of ${TLM}_b(A)_{j}$ over the $K = 50$ repetitions. The results are only given for the models using Tukey's M-estimators since, in general, these models performed the best on average. This is attributed to the corresponding $\psi$-function used in the M-estimation which trims large outliers from the data completely. Huber's version down-weights outliers, but never completely trims them.  The states are ordered along the $x$-axis according to number of agencies within the state (shown in parentheses). In several of the smaller states, the restricted hierarchical model performs better on average with similar performance between the models in most of the larger states; a reflection of the decreased influence of the prior. It is clear that the prior is having the effect of pooling information across states; beneficial for reducing the sample-to-sample variance. Figure \ref{fig:hierTLMstateSD} shows this benefit by plotting the standard deviation of ${TLM}_b(A)_{j}$ relative to the mean for each state. The general trend in the figure is that the standard deviation decreases as $n_{j}$ increases. Additionally, the differences between the models is larger for smaller states with virtually identical standard deviations for the larger states. Similar benefits are often seen for hierarchical models \citep[e.g.,][]{gelman2006}.

@@

\begin{figure}[t]
\includegraphics[width=7in]{hier_sd_tlm_state.png}
\caption{Hierarchical model results: Standard deviation of ${TLM}_b(A)_{j}$  relative to the mean over $K = 50$ repetitions for each state and $\alpha = 0.3$. The states are ordered along the $x$-axis according to number of agencies within the state (shown in parentheses). Results displayed are for the robust models using Tukey's M-estimators.}
\label{fig:hierTLMstateSD}
\end{figure}

>>>>

To complete the specification of model \eqref{fullsim2},  $a_{s}$ and $b_{s}$ are fixed to a variety of values representing different levels of prior knowledge. For each we set $b_{s} = 4a_{s}c$ resulting in a prior mean for each $\sigma^{2}_{i}$ of $\frac{4ca_{s}}{a_{s}-1}, \ a_{s} >1$. The precision is $\frac{(a_{s} -1)^{2}(a_{s}-2)}{(4ca_{s})^{2}}$; meaning larger $a_{s}$ and smaller $c$ result in a more informative prior. With $c = 1$ the shrinkage (for large $a_{s}$) is to the true value of $\sigma^{2} = 4$. We consider $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$ for a total of nine different priors whose densities are displayed in Figure \ref{fig:priors}. The vertical dashed line is at the known true value of $\sigma^{2} = 4$.

\begin{figure}[t]
\centering
\includegraphics[width = 4in]{priors_sigma2.png}
\label{fig:priors}
\caption{The nine different inverse gamma priors used for the group level variance parameters $\sigma_{j}^{2}$ in the simulation. The shape parameter values are $a_{s} = 1.25, 5, 10$. The scale parameter values are $b_{s} = 4a_{s}c$ with $c = 0.5, 1, 2$. The vertical dashed line in each panel is at the true value of $\sigma^{2} = 4$ and is the variance of the good portion of the data in the simulation.}
\end{figure}

@@

To assess predictive capability, the models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution. Specifically, for the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
\begin{equation}
\label{kl}
KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
\end{equation}
where $M$ indexes the fitting model, $f(\tilde y | \theta_{i}, \sigma^{2}) = N(\tilde y | \theta_{i}, \sigma^{2})$; the normal density function with mean $\theta_{i}$ and variance $\sigma^{2}$, evaluated at $\tilde y$. For the Bayesian models ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})d\theta_{i}d\sigma^{2}_{i}$ where $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions, along with specified $a_{s}$ and $c$. For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k}) = N(\tilde y|\hat\theta_{i}, \hat\sigma^{2}_{i})$ as a groupwise plug-in estimator for the predictive distribution. The classical fits are done separately for each group with no consideration of the hierarchical structure between the groups. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models, where smaller means correspond to better fits. Sampling variation is summarized with the standard error between the $K=30$ replicates in the simulation:  $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{1}{K(K-1)}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ where $\overline{KL}^{(M)}_{{\cdot} k} = \frac{1}{90}\sum_{i = 1}^{90} KL^{(M)}_{ik}$.

Figure \ref{kl_sim} displays $\overline{KL}^{(M)}_{{\cdot}{\cdot}}$  with error-bars plus/minus one $SE(\overline{KL}^{(M)}_{{\cdot} k})$ for each $a_{s} = 1.25,  5, 10$ and $c = 0.5, 1, 2$. The values of $a_{s}$ and $c$, do not effect the classical robust linear models. The average KL  for the normal theory models range between $0.22$ and $0.3$ which is much worse than the robust methods and hence are left out of the figure. For $c = 0.5$ and $c = 1$, the results are in favor of the restricted likelihood methods with perhaps a slight advantage to the use of Tukey's location estimator over Huber's. This is likely due to the fact that Tukey's estimator trims extreme outliers completely in the estimation procedure \citep{huber2009}. For $c = 2$, the classical methods win, again with a slight advantage to Tukey's statistic. It is clear that the priors picture in Figure \ref{fig:priors} matter. For $c = 2$, the priors subsequently add more mass significantly above $\sigma^{2} = 4$ as $a_{s}$ gets larger. Additionally the tuning parameters chosen for the location and scale estimators result in an upward bias in the estimate of $\sigma^{2}$. This bias depends on $m$ and $p$. For example, for $m = 9$ and $p = .1$, Huber's version converges to roughly $4.8$ asymptotically with $n$ under the data generating model. This bias gets worse for more severe contamination levels in the simulation. For $c = 0.5$ and $c = 1$, the priors place more mass on smaller values as $a_{s}$ grows. The larger values of $a_{s}$ also result in  better (smaller) average KL divergence.  For $c = 1$ the prior concentration is roughly around the true $\sigma^{2} = 4$. For $c = 0.5$ the prior concentration is roughly around half $\sigma^{2} = 2$. Interestingly, this prior performs better in terms of KL divergence than having the seemingly more correct prior with $c = 1$. This prior does not capture good prior information on $\sigma^{2}$, but does compensates more for the upward bias in the estimates of scale. This highlights one potential effect of the choice of statistic and associated tuning parameters when applying the restricted likelihood.

@@

It is also interesting to consider the effects of factors $n$, $p$, and $m$.  We subset the results to a single prior ($a_s = 5$ and $c = 1$) to not confound the effects of the different priors. For each simulation $k$, the main effect averages of $KL^{(M)}_{ik}$  are found for each factor $n$, $p$, and $m$. Figure \ref{kl_mnp} displays the average of these main effects over the $K = 30$ simulations along  with error bars plus/minus one standard error. For each group $n$, $p$, and $m$ the restricted versions have better (lower) average KL divergences. The patterns also makes sense: average KL with get larger (worse) as the contamination gets more severe (larger $m$ or larger $p$) and average KL gets smaller (better) as the sample size $n$ grows.

\begin{figure}[t]
\centering
\includegraphics[width = 4in]{kl_sim2_mnp.png}
\caption{Average KL - divergence plus/minus one standard error grouped by the factors $m$ (left), $n$ (middle), and $p$ (right). Results for a single prior ($a_s = 5$ and $c = 1$) are displayed to not confound the effects of the different priors.}
\label{kl_mnp}
\end{figure}

This simulation shows the potential of the restricted likelihood while highlighting some interesting observations. Specifically, the choice of summary statistics, along with corresponding tuning parameters is important. For the tuning parameters, we applied the default choice of $95\%$ efficiency at the normal.  Under the simulation model here, this choice results in bias in the scale estimation which affects the performance of the method. These choices must be made when using both the classical and Bayesian methods. The Bayesian setting allows for the incorporation of informative prior information which, as shown in this example, can improve prediction if chosen wisely. %In this severe outlier contamination setting, the better performance of the priors that shrink towards too small of values for $\sigma^{2}$ ($c = 0.5$) is an artifact of the fact that the tuning parameters on the estimators were non-optimal for this setting and more shrinkage was needed. However, a good choice for a prior on $\sigma^{2}$ ($c = 1$) still performs well. %A more refined model could have different tuning parameters for the groups representing different levels of contamination. However, a prior knowledge for

%The results are sensitive to the prior, but here we have observed good relative improvement over the classical counterparts for a range of prior choices.

@@

It is also interesting to examine the results within each state. Figure \ref{fig:hierTLMstate} summarizes ${TLM}_b(A)_{j}$ for each state where the points and errorbars are the averages and plus/minus one standard deviation of ${TLM}_b(A)_{j}$ over the $K = 50$ repetitions. The results are only given for the models using Tukey's M-estimators since, in general, these models performed the best on average. This is attributed to the corresponding $\psi$-function used in the M-estimation which trims large outliers from the data completely. Huber's version down-weights outliers, but never completely trims them.  The states are ordered along the $x$-axis according to number of agencies within the state (shown in parentheses). In several of the smaller states, the restricted hierarchical model performs better on average with similar performance between the models in most of the larger states; a reflection of the decreased influence of the prior. It is clear that the prior is having the effect of pooling information across states; beneficial for reducing the sample-to-sample variance. Generally,  the standard deviations are smaller for the hierarchical model for smaller states with virtually identical standard deviations for the larger states. Similar benefits are often seen for hierarchical models \citep[e.g.,][]{gelman2006}.

@@

%\begin{figure}[t]
%\includegraphics[width=7in]{hier_sd_tlm_state.png}
%\caption{Hierarchical model results: Standard deviation of ${TLM}_b(A)_{j}$  relative to the mean over $K = 50$ repetitions for each state and $\alpha = 0.3$. The states are ordered along the $x$-axis according to number of agencies within the state (shown in parentheses). Results displayed are for the robust models using Tukey's M-estimators.}
%\label{fig:hierTLMstateSD}
%\end{figure}
